\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts} 
% or 
\usepackage{amssymb}

\usepackage{mathtools}

\usepackage{tikz}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}
\usetikzlibrary{patterns}

\pgfplotsset{compat = newest}
\usepackage{cancel}
\usepackage[margin=1in]{geometry}

\begin{document}

\title{Maths Lessons}
\maketitle

\section{Quadratics}
Quadratics are second degree polynomials. They usually contain an $x^2$ and can be written
in the form $ax^2 + bx + c$.
\\
The 4 main methods of solving quadratics are as follows:
\begin{itemize}
	\item Factorise
	\item Quadratic Formula
	\item Complete the Square
	\item Graphically
\end{itemize}

\subsection{Quadratic Formula}
The quadratic formula is used to find the solution to arbitrary quadratics in the form $ax^2 + bx + c$. It is as follows:

\begin{equation}
	\frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
\end{equation}

\subsubsection{Deriving the Quadratic fromula}
The quadratic fromula is derived from completing the square. Completing the square is the first step
\begin{gather*}
	ax^2 + bx + c = 0\\
	a \left (x + \frac{b}{2a} \right )^2 - a\left (\frac{b}{2a} \right )^2 + c = 0 \\
	a \left (x + \frac{b}{2a} \right )^2 - \frac{b^2}{4a} + c = 0 \\
	a \left (x + \frac{b}{2a} \right )^2 = \frac{b^2}{4a} - c \\
	\left (x + \frac{b}{2a} \right )^2 = \frac{b^2}{4a^2} - \frac{c}{a} \\
	\left (x + \frac{b}{2a} \right )^2 = \frac{b^2 - 4ac}{4a^2} \\
	x + \frac{b}{2a} = \pm \sqrt{\frac{b^2 - 4ac}{4a^2}} \\
	x + \frac{b}{2a} = \frac{\pm \sqrt{b^2 - 4ac}}{2a} \\
	x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a} \\
\end{gather*}

\subsubsection{Discriminant}
The discriminant is part of the quadratic formula that can tell us various properties of the parabola.
The discriminant is as follows:
\begin{equation}
	D = b^2 - 4ac
\end{equation}
When the discriminant is positive we have two solutions:
\begin{equation}
	D > 0
\end{equation}
When the discriminant is equal to $0$ we have one solution:
\begin{equation}
	D = 0
\end{equation}
When the discriminant is negative we have no real solutions:
\begin{equation}
	D < 0
\end{equation}
Instead we have complex solutions in the form $ai$ where $i = \sqrt{-1}$.

\subsection{Solving Hidden Quadratics by Substitution}

Given a quadratic such as $2^{2n} + 2 \left ( 4^{n} \right ) + 1$ we can solve by putting in the form
$ax^2 + bx + c$.

\begin{gather}
	2^{2n} + 2 \left ( 2^{n} \right ) + 1 = 0 \\
	\left ( 2^n \right ) ^2 + 2 \left (2^n \right ) + 1 = 0 \\
\end{gather}

\section{Trig}

\begin{equation}
	\sec \theta \equiv \frac{1}{\sin \theta}
\end{equation}


\begin{equation}
	\mathrm{cosec} \theta \equiv \frac{1}{\cos \theta}
\end{equation}

\begin{equation}
	\cot \theta \equiv \frac{1}{\tan \theta}
\end{equation}

\begin{equation}
	\sin^2 \theta  + \cos^2 \theta \equiv 1
\end{equation}

\begin{equation}
	\tan^2 \theta + 1 \equiv \sec^2 \theta
\end{equation}

\begin{equation}
	1 + \cot^2 \theta \equiv \mathrm{cosec}^2 x
\end{equation}

\section{Calculus}

Differentiation, finding a function for the gradient.

\begin{gather*}
	f(x)\\
	m = \frac{y - y_1}{x - x_1}\\
	h = x - x_1 \\
	\text{We want to make $h$ as small as possible so we say that: }\\
	h \to 0 \\
	y = f(x + h) \\
	y_1 = f(x) \\
	f'(x) = \lim_{h \to 0} \frac{f(x + h) - f(x)}{h}
\end{gather*}

\begin{gather*}
	f(x) = x^2 \\
	f'(x) = \lim_{h \to 0} \frac{f(x + h) - f(x)}{h} \\
	f'(x) = \lim_{h \to 0} \frac{(x + h)^2 - x^2}{h} \\
	f'(x) = \lim_{h \to 0} \frac{x^2 + 2hx + h^2 - x^2}{h} \\
	f'(x) = \lim_{h \to 0} \frac{2hx + h^2}{h} \\
	f'(x) = \lim_{h \to 0} 2x + h \\
	f'(x) = 2x \\
\end{gather*}
Integration, finding a function for the area under a curve.
This is the opposite of differentiation so we can reverse it.

\subsection{Sum rule}

\begin{equation}
	(f + g)' = f' + g'
\end{equation}

\subsection{Product Rule}

\begin{equation}
	(fg)' = fg' + f'g
\end{equation}

\subsection{Chain Rule}

\begin{gather}
	h(x) = f(g(x))\\
	h'(x) = f'(g(x)) \cdot g'(x) 
\end{gather}

\subsection{Second derivative}

The second derivative finds the gradient of the gradient function. The notations $f''(x)$ and $\frac{d^2x}{dy^2}$ are commonly used.

\subsubsection{Deriving the notation}

\begin{gather}
	\frac{dy}{dx} \\
	\frac{d}{dx} \left ( \frac{dy}{dx} \right )\\
	\frac{d(dy)}{dx^2} \\
	\frac{d^2y}{dx^2} \\
\end{gather}

\subsubsection{Stationary Points}
Whenever $f'(x) = 0$, there is a stationary point.

\begin{align*}
	\text{Inflection:} &\left \{ f''(x) = 0 \right \} \\
	\text{Maximum:} &\left \{ f''(x) < 0 \right \} \\
	\text{Minimum:} &\left \{ f''(x) > 0 \right \} \\
\end{align*}

\subsection{Integration}

\subsubsection{Definite Integrals}

\begin{equation}
	\int x^n = \frac{x^{n + 1}}{n+1} + C
\end{equation}

\subsubsection{Indefinite Integrals}

\begin{equation}
	\int_a^b x^n = \left [ \frac{x^{n+1}}{n+1}\right ]_a^b= \frac{b^{n + 1}}{n+1} - \frac{a^{n+1}}{n+1}
\end{equation}

\section{Matrices}

\begin{gather*}
	\begin{pmatrix}
		a & b \\
		c & d \\
	\end{pmatrix} \cdot \begin{pmatrix}
		e & f \\
		g & h \\
	\end{pmatrix} = \begin{pmatrix}
		a \times e + b \times g & a \times f + b \times h \\
		c \times e + d \times g & c \times f + d \times h \\
	\end{pmatrix}
\end{gather*}

\begin{gather*}
	m = \begin{pmatrix}
		a & b \\
		c & d \\
	\end{pmatrix}\\
	m \cdot m^{-1} = \begin{pmatrix}
		1 & 0 \\
		0 & 1 \\
	\end{pmatrix}\\
	m^{-1} = \begin{pmatrix}
		10 & 1 \\
		1 & 10 \\
	\end{pmatrix}
\end{gather*}


\section{Binomial Expansion}

To exapand a binomial in the form $(a + b)^n$, we can use pascals triangle to work out the coeficients and exapnd from there.
For example:
\begin{gather*}
	(a + b)^3 \\
	\text{We use the 4th layer of the pascal triangle $1, 4, 4, 1$} \\
	a^3 + 4a^2b + 4ab^2 + b^3 \\
\end{gather*}
To write a more general formula for this we need to remind ourselves of factorials:
\begin{equation}
	n! = n \times (n - 1) \times (n - 2) \times \cdots \times 1
\end{equation}
This tells us combinations for ordering things. E.g, if given $4$ books they can be put in $4!$ different
orders. We might want to choose a set of $r$ values from a set of $n$ numbers. The function that gives us the number of
combinations for this is as follows.
\begin{equation}
	\binom{n}{r} = \prescript{n}{}{\mathbf{C}_r} = \frac{n!}{(n - r)!r!}
\end{equation}
Using this we can define an expression for arbitrary binomial expansion as follows:
\begin{equation}
	(a + b)^n = a^n + \binom{n}{1}a^{n - 1}b + \binom{n}{2}a^{n - 2}b^{2} + \cdots + \binom{n}{r}a^{n - r}b^{r} + \cdots + b^n
\end{equation}
We can also define an expression for the term at index $r$ in a binomial expansion:
\begin{equation}
	\binom{n}{r}a^{n - r}b^{r}
\end{equation}

\section{Linear Graphs}
Linear graphs are graphs in the form $y = mx + c$ and produce a straight line.

\subsection{Paralel and Perpendicular}
Two linear graphs are parallel if their gradients are equal.
\begin{equation}
	m_1 = m_2
\end{equation}
Two linear graphs are perpendicular if their gradents are negative recipricles of eachother. We can express this rule as the following

\begin{equation}
	m_1 \times m_2 = -1
\end{equation}
This works because the any number multiplied by it's reciprical is always $1$

\begin{equation}
	a \times a^{-1} = 1
\end{equation}

\subsection{Suplimentry rules}

\subsubsection{Calculating the Gradient}

\begin{equation}
	m = \frac{y_1 - y_2}{x_1 - x_2}
\end{equation}

\subsubsection{Easier Offset}

\begin{equation}
	y - y_1 = m(x - x_1)
\end{equation}

\subsubsection{Complete Form}

\begin{equation}
	y - y_1 = \frac{y_1 - y_2}{x_1 - x_2}(x - x_1)
\end{equation}
From there we can substitute in values for $x_1$, $x_2$, $y_1$ and $y_2$ and rearrange into the required form
(typically $y=mx+c$ or $ax + by + c = 0$).

\section{Circles}

\subsection{Circle Theorems}

\begin{center}

	\begin{tikzpicture}
		\coordinate [label={below right:$C$}] (C) at (0, 0);
		\fill [black] (C) circle (0.075);
		\draw (C) circle [radius=3]; 
		\draw (-3,-3) -- (3,3);
		\fill [black] (1.5,1.5) circle (0.075);
		\fill [black] (0,3) circle (0.075);
		\fill [black] (3,0) circle (0.075);
		\draw (0,3) -- (3,0);

	\end{tikzpicture}

\end{center}

\section{asymptotes}
Asymptotes are imaginary lines that graphs will extend to but never touch. A good example of this is with the graph $y=x^{-1}$
which has two asymptotes. One on the x axis ($y = 0$) and one on the y axis ($x = 0$)

\begin{center}
	\begin{tikzpicture}
		\begin{axis}[
			xmin=-10, xmax=10,
    		ymin=-10, ymax=10,
			axis lines=middle
		]
			\addplot[samples=100,domain=-10:10] {x^(-1)};
		\end{axis}
	\end{tikzpicture}
\end{center}

\section{Exponentials and Logarythms}

\subsection{Laws of Logarythms}

\begin{gather}
	\log_{a}{x^y} = y\\
	\log_{a}{x} + \log_{a}{y} = \log_{a}{xy} \\
	\log_{a}{x} - \log_{a}{y} = \log_{a}{\frac{x}{y}} \\
	y\log_{a}{x} = \log_{a}{x^y}
\end{gather}

\subsubsection{Special Cases}

\begin{gather}
	\log_{a}{\frac{1}{x}} = \log_{a}{x^{-1}} = -\log_{a}{x} \\
	\log_{a}{a} = 1 \left \{a > 0, a \ne 1 \right \}\\
	\log_{a}{1} = 0 \left \{a > 0, a \ne 1 \right \}\\
\end{gather}

\subsection{Natural log}

\begin{gather}
	\ln{x} = \log_e{x}
\end{gather}

\subsection{Graphs}

Both $ln x$ and $e^x$ have one asymptote. In the case of $ln x$ it is the y axis ($x = 0$) and for $e^x$ it is the x axis ($y = 0$).
They are symetrical along the line $y=x$. The graphs have been plotted below:

\begin{center}
	\begin{tikzpicture}
		\begin{axis}[
			xmin=-10, xmax=10,
    		ymin=-10, ymax=10,
			axis lines=middle,
			tick style={draw=none},
			xticklabels={,,},
			yticklabels={,,}
		]
			\addplot[samples=100, domain=-10:3] {exp(x))};
			\addplot[samples=100, domain=0:10] {ln(x))};
			\addplot[dashed] {x};
		\end{axis}
	\end{tikzpicture}
\end{center}

\section{Tangents and Normals}
Normals are lines perpendicular to a surface where as tangents are lines
that touch the surface at a single point with the same gradient.

\begin{center}
	\begin{tikzpicture}
		\begin{axis}[
			xmin=-10, xmax=10,
    		ymin=-10, ymax=10,
			axis lines=middle
		]
			\addplot[samples=100,domain=-10:10] {x^2};
			\addplot[samples=100,domain=-10:10] {2 * 1 * x - 1^2};
			\addplot[samples=100,domain=-10:10] {1 - ((x - 1) / (2 * 1))};
		\end{axis}
	\end{tikzpicture}
\end{center}
The gradients of the tangents and normals have the following relationship
\begin{equation}
	m_n = - m_t^{-1}
\end{equation}
\\
Arbitrary tangent and normal functions ($f_t(x, a)$ and $f_n(x, a)$) of a given function $f(x)$ at a given offset $a$ can be given as follows:

\begin{gather}
	f_t(x, a) = f'(x)(x-a) + a^2 \\
	f_n(x, a) = a^2 -f'(x)^{-1}(x-a) \\
\end{gather}

\section{Factor theorem}
When given a polynomial $f(x) = ax^n + bx^{n-1} + \cdots + z $
An expression $(x + a)$ is a factor of $f(x)$ if $f(-a) = 0$.
\\
We can use this to rule to factorise higher order polynomials such as the following:

\begin{gather}
	f(x) = x^3 + 3x^2 + 3x + 1 \\
	f(-1) = (-1)^3 + 3(-1)^2 + 3(-1) + 1 = 0\\
	\therefore (x + 1) \text{ is a factor} \\
	f(x) = (x+1)b = x^3 + 3x^2 + 3x + 1
\end{gather}

\begin{center}
	\begin{tabular}{ c | c | c | c }
		$\times$ & $x^2$ & $2x$ & $1$\\ \hline
		$x$ & $x^3$ & $2x^2$ & $x$  \\ \hline
		$1$ & $x^2$ & $2x$ & $1$\\
	\end{tabular}
\end{center}

\begin{gather}
	f(x) = (x+1)(x^2 + 2x + 1) \\
	f(x) = (x+1)(x+1)(x+1) = (x+1)^3
\end{gather}

\section{Forces}
Forces are vectors and are made from a magnitude and direction.

Magnitude can be written as the following
\begin{equation}
	\lvert F \rvert
\end{equation}

Direction measured anticlockwise from the horrizontal.

can be represented in terms of $i$ and $j$

\subsection{Labelling Forces}

Normal reaction is perpendicular to surface.
\textbf{TODO: TIKZ}

Weight acts vertical.

Tension acts on a string

friction always opposes motion
thrust or compression 



\subsection{Assumptoins}
\begin{itemize}
	\item We assume there are no extenral forces such as friction or air resistance or wind
	\item We assume things are particles
	\item gravity is constant
	\item we assume rigidity
	\item we assume friction for a block sliding down a hill
	\item See slides for more detail.
\end{itemize}



Resultant force is the sum of all of the forces acting on an object. It is the actual force that
happens to the object

\begin{equation}
	R = \sum F
\end{equation}


Euqalibrium means net force is $R = 0$

\section{Newtons Laws}

\subsection{Newtons First Law}
\begin{equation}
	\sum F = 0
\end{equation}
The body remains at rest at a state of uniform motion unless the forces acting on it are not in equlirium
\\\\
An object will stay still or carry on moving exactly the way it was unless unbalanced force acts on it.
\\\\
When there is no resultant force (it is at an equalibriumdc, the object remains stationary or at a constant velocity

\subsection{Newtons Second Law}
\begin{equation}
	\sum F \ne 0
\end{equation}
When there is a resultant for the object accelerates, changes direction or deformation (change in structure).

\begin{equation}
	\sum F = ma
\end{equation}
If an unbalanced force acts on an object then its velocity will change.
It will either change direction, speed up or slow down, and that includes stopping.

\subsection{Newtons Third Law}
Every action has an equal and oposite reaction \ldots \textbf{of the same type}.
\\\\
Compare the same forces, can't compare normal contact force and weight.
\\\\
When body A exerts a force on body B, an equal an oposite force \textbf{of the same type} is exerted by
body B on body A.
\\\\
Normal contact force is electrostatic force as you don't technically touch anything.


\section{Non-Uniform Acceleration}

\begin{equation}
	a = f(t)
\end{equation}


When acceleration isn't constant displayment can be given by any function; not just a quadratic.

\begin{gather}
	v(t) = s'(t) \\
	a(t) = v'(t) = s''(t) \\
	v = \frac{ds}{dt} \\
	a = \frac{dv}{dt} = \frac{d^2s}{dt}
\end{gather}


when  $s = f(t)$ then at a time of $4$ for example, $s = f(4)$
To calculate velocoty from this we differentiate
Initial velocity is when $t = 0$


\begin{gather}
	s(t) = \int v(t) dt = \iint a(t) dt^2 \\
	v(t) = \int a(t) dt
\end{gather}

Initial conditions. there will always be a constant. \textbf{TODO: LOOK UP ABOUT THIS}

\section{Partial Fractions}

\begin{gather}
	\frac{ax^2+bx+c}{(x+e)(x+f)} = A + \frac{B}{x+e} + \frac{C}{x+f}\\
	ax^2+bx+c = A(x+e)(x+f) + B(x+f) + C(x+e)\\
	\text{Then substitute $-e$ and $-f$ to find $A$, $B$ or $C$} \\
	\text{Or Equate Coefficients:} \\
	a = A \\
	b = A(e + f) + B + C \\
	c = Aef + Be + Cf 
\end{gather}

\subsection{Repeated Powers}

\begin{equation}
	\frac{ax + b}{(x + c)^2} = \frac{A}{x+c} + \frac{B}{(x+c)^2}
\end{equation}


\section{Integration by subsitution}
stuffs

\begin{gather}
	\int \frac{1}{x+2} dx \\
	\text{let } p = x + 2 \\
	\frac{dp}{dx} = 1 \\
	\frac{dp}{1} = dx \\
	\int \frac{1}{p}dp = \ln{\left |p \right | + C} \\
	\ln \left |x + 2 \right | + C
\end{gather}

\section{Compound Angle Formule}

\begin{gather}
	\sin ( A \pm B ) \equiv \sin A \cos B \pm \cos A \sin B \\
	\cos ( A \pm B ) \equiv \cos A \cos B \mp \sin A \sin B \\
	\tan ( A \pm B ) \equiv \frac{\tan A \pm \tan B}{1 \mp \tan A \tan B} \\
	\tan ( A \pm B ) \equiv \frac{\sin(A \pm B)}{\cos (A \pm B)}
\end{gather}

\subsection{Double Angle Formule}

\begin{gather}
	\sin 2A \equiv 2 \sin A \cos A \\
	\cos 2A \equiv \cos^2 A - \sin^2 A \\
	\cos 2A \equiv 2\cos^2 A - (1 - \cos^2 A) \\
	\cos 2A \equiv (1 - \sin^2 A) - \sin^2 A \\
	\tan 2A \equiv \frac{2 \tan A}{1 - \tan^2 A}
\end{gather}

\subsection{Harmonic Form}

\begin{gather}
	a \sin \theta \pm b \cos \theta \equiv R \sin (\theta \pm \alpha) \\
	a \cos \theta \pm b \sin \theta \equiv R \cos (\theta \mp \alpha)
\end{gather}

\begin{gather}
	2 \sin x + 5 \cos x \equiv R \sin (x + \alpha)\\
	2 \sin x + 5 \cos x \equiv R \sin x \cos \alpha + R \cos x \sin \alpha \\
	R \cos \alpha = 2 \\
	R \sin \alpha = 5 \\
	\tan \alpha = \frac{R \sin \alpha}{R \cos \alpha} = \frac{5}{2} \\
	(R \sin \alpha)^2 + (R \cos a)^2 \equiv R^2(\sin^2 \alpha + \cos ^2 \alpha) \equiv R^2 \\
	\because \sin^2 \alpha + \cos^2 \alpha \equiv 1
\end{gather}

\end{document}
